Supplementary Materials for: The promise of community-driven preprints in ecology and evolution

### 2024-09-25 supp/supp.qmd
Update supp and remove unecessary csl

---
title: "Supplementary Materials for: Empowering the use of community driven preprints in ecology and evolution"
bibliography: ../bib/refs.bib
csl: ../bib/nature.csl
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "**Table of Contents**"
    theme: simplex
    embed-resources: true
    code-fold: show
    code-tools: true
    number-sections: true
    fontsize: "12"
    max-width: "10"
    code-overflow: wrap
  docx:
    reference-doc: ../bib/template.docx
  pdf:
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'Figure'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "Figure"
  tbl-prefix: "Table"
---

@@

We had a total of `r del_details[2,2]` participants in the workshop, and an additional `r del_details[1,2]` that could not attend but were interested in contributing. Participants assigned themselves to collect data from between `r range(contr_extract$n)[1]` to `r range(contr_extract$n)[2]` preprints (mean = `r mean(contr_extract$n)`, standard deviation = `r sd(contr_extract$n)`). Data collection was achieved using a Google Form to standardise data input. Participants collected the following information for each of their assigned preprints:

@@

- **Country of the corresponding / submitting author**: Country of affiliation for the author submitting the preprint collecting accoring to standard ISO 3166 names. If teh country could not be determined the affiliation was considered as "NA". 
- **Year of first publication for corresponding/submitted author**: We used [Google Scholar](https://scholar.google.com.au) to collect publication year of the first journal article published by author submitting the preprint. If the submitting author had no profile or did not publish the value enetred was '0'.

@@

- **Type of article**: *EcoEvoRxiv* publishes a greater diversity of preprints compared with other preprint servers. We therefore collected data on article type. The categories were: *Research Article*: any article-like manuscript intended for publication in research journals with new empirical findings; *Methods paper*: papers presenting new methodological or computational approaches; *Reviews and Meta-analyses*: papers quantitatively or qualitatively synthesising a given topic;  *Opinions*: usually short papers providing new perspectives on a topic; *Comments*: papers that explicitly comment on an already published research article; and *Other*: which includes any other category of preprint which may also be government, non-profit, or industry reports, white papers, or other documents that are not intended for publication in a research journal.
- **Link to Data for Preprint**: We were explicitly interested in whether data for the preprint was already publically available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying data was provided. If it was reported we took the link to the data.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository was reported we entered 'none'. If multiple links were provided we took one.
- **Link to Code for Preprint**: We were also interested in whether code for data analysis was already publically available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying code was provided. If it was reported we took the link to the code.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository with code was reported we entered 'none'. If multiple links were provided we took one.
- **Number of citations to preprint**: To understand citation patterns of preprints on *EcoEvoRxiv* we used Google Sscholar to search for the number of citations attributed to the preprint. If the preprint version was not on Google Scholar or was merged with an already published version we considered it as 'NA' (coded as '999'). If the preprint version was online and clearly indicated (i.e., EcoEvoRxiv) then we took the number of citations at the time of collection. While citation counts may vary slightly, the first round of data collection was done in a short time frame. As such, we do not anticipate that citation counts would have changed significantly between the time of data collection and the time of submission of this manuscript.
- **PCI recommendation**: We also deteremined if the preprint been recomended by Peer Community In (PCI). To do this, we searched for any PCI recommendations associated with the preprint on the preprint landing page. If that none was available on the landing page then we searched "peercommunityin.org recommendation "TITLE OF PREPRINT"" on Google. If a link was discovered, we indicated 'yes' and took the DOI link. If not, then we assumed it was not recommended.

@@

- **Link to Data for Publication**: If the preprint was published we collected the link to any published data provided on the publishers webpage or within the published paper. If a link to the data was not visble we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. 
- **Link to Code for Publication**: If the preprint was published we collected the link to any published code provided on the publishers webpage or within the published paper. If a link to the data was not visble we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'.

Additional variables of interest (i.e., journal impact factor, preprint and review policy, gender of submitting author) could be determined oustide of this data collection period.

## Data Checking and Validation
Data from each preprint was cross checked by two authors to ensure accuracy and consistency of the data collected. Any discrepencies were either discussed or fixed prior to analysis.

@@

#### Generational Analysis
To test whether a greater number of preprints were submitted by authors of earlier 'academic age', as quantified by the submitting authors first year of publication, we used a generalised linear model (GLM) with a negative binomial error distribution. We included the number of preprints submitted in a given year as the response variable and the first year of publication as the predictor variable. A negative, and significant, slope coefficient indicates support for researchers of earlier career stages submitting more preprints.

@@

Interestingly, a highly cited preprint did not necessarily become a highly cited published article over time, although some of this descrepency is related to whether and how preprints and articles are linked through crossref (Spearman Rank Correlation: `r spear$estimate`, S = `r spear$statistic`, *p* = `r spear$p.value`; @fig-citations).

>>>>

---
title: "Supplementary Materials for: The promise of community-driven preprints in ecology and evolution"
bibliography: ../bib/refs.bib
csl: ../bib/proceedings-of-the-royal-society-b.csl
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    toc-title: "**Table of Contents**"
    theme: simplex
    embed-resources: true
    code-fold: show
    code-tools: true
    number-sections: true
    fontsize: "12"
    max-width: "10"
    code-overflow: wrap
  docx:
    reference-doc: ../bib/template.docx
  pdf:
    include-in-header: 
      text: '\usepackage{lineno}\linenumbers'
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto  # re-render only when source changes
  cache: false
  echo: false
  warning: false
  error: false
  include: true
crossref:
  fig-title: 'FigureS'
  fig-labels: arabic
  title-delim: "-"
  fig-prefix: "FigureS"
  tbl-prefix: "TableS"
---

@@

We had a total of `r del_details[2,2]` participants in the workshop, and an additional `r del_details[1,2]` that could not attend but were interested in contributing. Participants assigned themselves to collect data from between `r range(contr_extract$n)[1]` to `r range(contr_extract$n)[2]` preprints (mean = `r mean(contr_extract$n)`, standard deviation = `r sd(contr_extract$n)`). Data collection was achieved using a Google Form to standardize data input. Participants collected the following information for each of their assigned preprints:

@@

- **Country of the corresponding / submitting author**: Country of affiliation for the author submitting the preprint collecting according to standard ISO 3166 names. If teh country could not be determined the affiliation was considered as "NA". 
- **Year of first publication for corresponding/submitted author**: We used [Google Scholar](https://scholar.google.com.au) to collect publication year of the first journal article published by author submitting the preprint. If the submitting author had no profile or did not publish the value entered was '0'.

@@

- **Type of article**: *EcoEvoRxiv* publishes a greater diversity of preprints compared with other preprint servers. We therefore collected data on article type. The categories were: *Research Article*: any article-like manuscript intended for publication in research journals with new empirical findings; *Methods paper*: papers presenting new methodological or computational approaches; *Reviews and Meta-analyses*: papers quantitatively or qualitatively synthesizing a given topic;  *Opinions*: usually short papers providing new perspectives on a topic; *Comments*: papers that explicitly comment on an already published research article; and *Other*: which includes any other category of preprint which may also be government, non-profit, or industry reports, white papers, or other documents that are not intended for publication in a research journal.
- **Link to Data for Preprint**: We were explicitly interested in whether data for the preprint was already publicly available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying data was provided. If it was reported we took the link to the data.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository was reported we entered 'none'. If multiple links were provided we took one.
- **Link to Code for Preprint**: We were also interested in whether code for data analysis was already publicly available. To collect these data we scrolled through the first version of the preprint and any related files on *EcoEvoRxiv* to see if a link (or accession number) to underlying code was provided. If it was reported we took the link to the code.  If the preprint was not an empirical or equivalent article based on data and data analyses (so usually for opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. If the preprint was based on data but no link (or accession number) to data repository with code was reported we entered 'none'. If multiple links were provided we took one.
- **Number of citations to preprint**: To understand citation patterns of preprints on *EcoEvoRxiv* we used Google Scholar to search for the number of citations attributed to the preprint. If the preprint version was not on Google Scholar or was merged with an already published version we considered it as 'NA' (coded as '999'). If the preprint version was online and clearly indicated (i.e., EcoEvoRxiv) then we took the number of citations at the time of collection. While citation counts may vary slightly, the first round of data collection was done in a short time frame. As such, we do not anticipate that citation counts would have changed significantly between the time of data collection and the time of submission of this manuscript.
- **PCI recommendation**: We also determined if the preprint been recommended by Peer Community In (PCI). To do this, we searched for any PCI recommendations associated with the preprint on the preprint landing page. If that none was available on the landing page then we searched "peercommunityin.org recommendation "TITLE OF PREPRINT"" on Google. If a link was discovered, we indicated 'yes' and took the DOI link. If not, then we assumed it was not recommended.

@@

- **Link to Data for Publication**: If the preprint was published we collected the link to any published data provided on the publishers webpage or within the published paper. If a link to the data was not visible we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'. 
- **Link to Code for Publication**: If the preprint was published we collected the link to any published code provided on the publishers webpage or within the published paper. If a link to the data was not visible we recorded the paper as not having provided data ('none'). If the paper had not been published or not based on data and data analyses (opinions, comments, narrative reviews, theoretical paper, etc.) we entered 'NA'.

Additional variables of interest (i.e., journal impact factor, preprint and review policy, gender of submitting author) could be determined outside of this data collection period.

## Data Checking and Validation
Data from each preprint was cross checked by two authors to ensure accuracy and consistency of the data collected. Any discrepancies were either discussed or fixed prior to analysis.

@@

#### Generational Analysis
To test whether a greater number of preprints were submitted by authors of earlier 'academic age', as quantified by the submitting authors first year of publication, we used a generalized linear model (GLM) with a negative binomial error distribution. We included the number of preprints submitted in a given year as the response variable and the first year of publication as the predictor variable. A negative, and significant, slope coefficient indicates support for researchers of earlier career stages submitting more preprints.

@@

Interestingly, a highly cited preprint did not necessarily become a highly cited published article over time, although some of this discrepancy is related to whether and how preprints and articles are linked through crossref (Spearman Rank Correlation: `r spear$estimate`, S = `r spear$statistic`, *p* = `r spear$p.value`; @fig-citations).
